---
title: "Appendix: BC Parks Attendance"
author: "Dayna K. Weststrate"
date: "2023-03-27"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)

# Import the model
M = readRDS("~/Desktop/bio/440/BCParks_Attendance/Scripts/models/model.rda")
# Import the climate projection data
climate = readRDS("~/Desktop/bio/440/BCParks_Attendance/Data/climate-projections/monthly-climate-projections.rds")
# Import park coordinates
park_coordinates <- read.csv("~/Desktop/bio/440/BCParks_Attendance/Data/bcparks/park_coordinates.csv")
# Import relevant population projection data (LG/HG)
LGrates <- read.csv('~/Desktop/bio/440/BCParks_Attendance/Data/projections/LG-rate-projections.csv')
HGrates <- read.csv('~/Desktop/bio/440/BCParks_Attendance/Data/projections/HG-rate-projections.csv')
# Import attendance projections
LGattendance = readRDS("~/Desktop/bio/440/BCParks_Attendance/Data/projections/LG-attendance-projections.rds")
HGattendance = readRDS("~/Desktop/bio/440/BCParks_Attendance/Data/projections/HG-attendance-projections.rds")
# Import historical attendance
bcparks <- readRDS("~/Desktop/bio/440/BCParks_Attendance/Data/bcparks/bcparks.rds")
# Remove camping data (unused in this study)
bcparks <- bcparks[!bcparks$attendancetype == "camping",] 
# Remove NA values
bcparks <- na.omit(bcparks)
```

# Overview

This appendix details the steps we used to analyze our data, build our model, project attendance, and visualize our results.

# Data Importing

Attendance records were obtained directly through BC Parks. There were monthly visitor totals for two attendance types; camping and day-use. Our project focused on day-use attendance, but an identical process could be executed using BC Parks camping data (or a combination of both attendance types).

The records included dates, visitor counts, and park names. Park names were assigned by BC Parks (`park` column). Each park was annotated with one of five regions; Northern, South Coast, West Coast, Thompson-Cariboo, or Kootenay-Okanagan (abbreviated in the `region` column as `north`, `south`, `west`, `tc`, and `ok`, respectively). Park region was obtained by searching each park on BC Parks' ArcGIS map and finding its classification under 'Park Region Name' (Esri, 2023). Parks belonging to Omineca Peace or North Coast Skeena regions were grouped into the Northern Region.

```{r package loading}
# Load necessary packages
library('readr')
library('stringi')
library('tidyr')        # for data wrangling
library('dplyr')        # for data wrangling
library('canadianmaps') # to download a shapefile of BC
library('elevatr')      # to download digital elevation models
library('purrr')        # for functional programming (map_***(), etc.)
library('sp')           # for spatial data
library('sf')           # for spatial data
library('terra')        # for raster data
library('lubridate')    # for formatting dates
library('ggplot2')      # for fancy plots
library('khroma')       # for fancy colour palettes
library('gridExtra')    # for arranging figure panels
library('lme4')         # for modelling
library('mgcv')         # for modelling
library('viridis')      # for fancy colour palettes
library('ggh4x')        # to fill in facet wrap title boxes
```

```{r data import, eval = FALSE}
# Import BC Parks attendance data
bcparks <- read.csv("~/Desktop/bio/440/BCParks_Attendance/Data/bcparks/dayuse.csv", 
                   na = "0", check.names = FALSE)
```


## Clean BC Parks Attendance Data

The raw BC Parks attendance data required wrangling. We converted the camping and day-use datasets from wide to long format, combined the data from both attendance types into one dataset, and created month and year columns from the dates.

```{r bcparks clean, eval = FALSE}
# Convert from wide to long format
bcparks <- gather(bcparks, date, visitortotal, "2010-01-01":"2019-12-01", 
                 factor_key = TRUE)
# Remove months with no values (NAs)
bcparks <- na.omit(bcparks)
# Classify type of attendance
bcparks$attendancetype <- "dayuse"

# Add new columns: Month and Year
bcparks <- bcparks %>%
  separate(date, sep="-", into = c("year", "month", "day"))
# Bring back the date column
bcparks$date <- paste(paste(bcparks$year, bcparks$month, sep = "-"), 
                      15, sep = "-") # include arbitrary day of the month
# Ensure correct formatting of columns
bcparks$date <- as.POSIXct(bcparks$date, format = "%Y-%m-%d")
bcparks$month <- as.numeric(bcparks$month)
bcparks$year <- as.numeric(bcparks$year) 
bcparks$park <- as.factor(bcparks$park)
# Remove day column (arbitrary number)
bcparks = subset(bcparks, select = -c(day))

```

# Data Annotation

## Population

Historical BC population data for 2010 through 2022 were annual estimates obtained from the Government of BC (2022) and were added under the `BCPop` column.

The `visitortotal` column represents visitation as calculated by BC Parks Operators. We corrected these counts for BC population by converting visitor totals into visitors per 1,000 BC residents (`attendance` column).

```{r population, eval = FALSE}
# Import historic population data
population_records <- read.csv("~/Desktop/bio/440/BCParks_Attendance/Data/population/population_records.csv")
# Remove unnecessary growth rate column
population_records = subset(population_records, select =-c(3))
# Add historic population as a column
bcparks <- merge(bcparks,population_records,by=c("year"))
# Correct visitor counts for population size
bcparks$attendance <- bcparks$visitortotal/bcparks$BCpop*1000
```

## Park Coordinates

The latitude and longitude coordinates of each park were determined by, again, using BC Parks' ArcGIS map to find park bounds, then placing a pin in Google Maps around the center of the park (Esri, 2023; Google, n.d.). Of course, this process is not exact, but the coordinates were used for downloading climate data proximate to each park, and weather is not likely to differ much throughout a given park.

```{r coordinates, eval = FALSE}
# Import park coordinate data
park_coordinates <- read.csv("~/Desktop/bio/440/BCParks_Attendance/Data/bcparks/park_coordinates.csv")
# Add latitude and longitude as columns
bcparks <- merge(bcparks,park_coordinates, by=c("park", "region"))
```

## Weather

Historical climate data were downloaded to attain average monthly temperature (ºC) and precipitation (mm) at each park (Wang et al., 2016). We cleaned the downloaded data so as to remove irrelevant variables and convert monthly total precipitation to average monthly precipitation. Average monthly temperatures and precipitation were represented by the columns `avgtemp` and `avgprecip`, respectively.

### Annotate coordinates with elevations

To download climate data from `climatenaR`, each coordinate required elevation, so we annotated the park coordinate CSV file with respective elevations.

```{r elevations, eval = FALSE}
# Convert all telemetry datasets to spatial data points
locations <- SpatialPoints(select(park_coordinates, longitude, latitude))
ctmm::projection(locations) <- '+proj=longlat'

# Import a Digital Elevation Model (DEM) for the region(s) of interest
dem <- get_elev_raster(locations = locations,
                       z = 3,
                       clip = 'bbox',
                       expand = 0.1)

# Write the CSV
park_coordinates <- mutate(locations_df,
                           el = terra::extract(dem, locations),
                           ID1 = 1:n(),
                           ID2 = ID1) %>%
  relocate(ID1:ID2)

write.csv(locations_df, file = 'Data/bcparks/parks-dem.csv', row.names = FALSE)
```

### Download and clean historical climate data

```{r historical climate, eval = FALSE}
library('climatenaR')   # to download climate data and projections

# If necessary, install the `climatenaR` package with
# `remotes::install_github('burnett-m/climatenaR', build_vignettes = TRUE)`
# *NOTE:* the `climatenaR` package requires the ClimateNA software.
# See https://register.climatena.ca/ to download it.

# Change the working directory as required by `climatenaR`
setwd('~/Desktop/bio/440/BCParks_Attendance/climate-na/')

# Download historical climate data
for(y in 2010:2022) {
  cat('Downloading ', y, '...\n', sep = '') # to track progress
  histClimateNA(
    file = 'parks-dem.csv',
    dateR = as.character(y),
    tFrame = 'M', # monthly averages
    exe = 'ClimateNA_v7.31.exe',
    outdir = 'dayna-s-park-climate-data') # exe location (in working directory)
}

# Clean historical climate data files
historicaldata <-
  # list all files, and import each of the CSVs
  map_dfr(
    list.files('Data/historical-climate', full.names = TRUE),
    \(.fname) {
      readr::read_csv(.fname, col_types = '?') %>%
        # add a column of the file name
        mutate(file = .fname)
    }) %>%
  mutate(year = substr(file,
                       start = stri_locate_first(file, regex = 'dem_')[1] + 4,
                       stop = nchar(file) - nchar('.csv'))) %>%
  # only keep relevant columns 
  select(year, Latitude, Longitude, Elevation, Tave01, Tave02, Tave03,
         Tave04, Tave05, Tave06, Tave07, Tave08, Tave09, Tave10, Tave11, Tave12,
         PPT01, PPT02, PPT03, PPT04, PPT05, PPT06, PPT07, PPT08, PPT09, PPT10,
         PPT11, PPT12) %>% 
  # pivot from wide to long format (only one column of precip and temp)
  pivot_longer(-c(year, Latitude, Longitude, Elevation),
               names_to = 'parameter', values_to = 'value') %>%
  # extract month and year out of parameters
  mutate(month = map_chr(parameter,
                         \(.chr) substr(.chr, nchar(.chr) - 1, nchar(.chr))),
         dec_date = decimal_date(date(paste(year, month, '15', sep = '-'))),
         month = as.numeric(month),
         year = as.numeric(year),
         parameter = map_chr(parameter,
                             \(.chr) substr(.chr, 1, nchar(.chr) - 2))) %>%
  # pivot wider to make separate columns of temperature and precipitation
  pivot_wider(names_from = parameter, values_from = value) %>%
  # convert monthly total precipitation to average monthly precipitation
  mutate(first_day = as.Date(paste(year, month, '01', sep = '-')),
         next_month = if_else(month != '12', as.numeric(month + 1), 1),
         next_year = if_else(month != '12', year, year + 1),
         last_day = as.Date(paste(next_year, next_month, '01', sep = '-')),
         samples = as.numeric((last_day - first_day)),
         avgprecip = PPT / samples) %>% # convert to millimeters per day
  # drop temporary columns
  select(-c(first_day, next_month, next_year, last_day, samples, PPT)) %>%
  # change column names
  rename(avgtemp = Tave,
         latitude = Latitude,
         longitude = Longitude,
         elevation = Elevation) %>%
  relocate(c(month, dec_date), .after = year) 
# Remove unnecessary columns from dataset
historicaldata = select(historicaldata, -c(dec_date, elevation))


# Add new columns to BC Parks dataset: Average temperature and precipitation 
bcparks <- merge(x=bcparks,y=historicaldata, 
                 by=c("latitude","longitude", "year", "month"), all.x=TRUE)
```

## Recreate Figure 1

### Figure 1A: Park Distribution Map

```{r fig1A}
theme_set(theme_map())

# Import a shapefile of British Columbia
bc_shp <-
  st_as_sf(PROV) %>% # convert to spatial features (sf) object
  filter(PRENAME == 'British Columbia') %>% # filter to BC only
  st_geometry() # extract boundaries only

# Filter dataset such that each park only has one observation 
parkcoords <- bcparks %>% distinct(park, latitude, longitude, .keep_all = TRUE)

# Plot parks on a map of BC
map <-
  ggplot() +
  ggtitle("A")+ # label the panel
  geom_sf(data = bc_shp) + # outline of BC
  geom_point(aes(longitude, latitude, col = region), parkcoords, 
             size = 3, shape = 17, alpha = 0.6) + # a single point for each park
  guides(col = guide_legend(override.aes = list(alpha=1))) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +
  theme(legend.title = element_text(size = 11, face = "bold"),
        legend.text = element_text(size = 11),
        legend.position = "bottom",
        legend.justification = "center",
        legend.direction = "horizontal",
        legend.box.background = element_rect(color = "black"),
        plot.margin = unit(c(-1,0,-1,0), "cm"),
        plot.title = element_text(vjust = -8.5, hjust = 0.03,
                                  size = 30, family = "sans", face = "bold")) +
  coord_sf() # ensures points don't get jittered around when figure dimensions change
map
```

### Figure 1B: Historical Seasonal Trends

```{r fig1B}
seasonal <-
  ggplot() +
  geom_jitter(data = bcparks,
              aes(y = attendance, x = month, col = region),
              alpha = 0.05, size = 1, width = 0.25,shape = 17) + # point for every observation
  geom_smooth(data = bcparks, 
              aes(y = attendance, x = month, col = region),
              size = 0.6, se = F) + # line for each region's trends
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  scale_y_continuous(limits = c(0, 15)) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +
  xlab("Month") +
  ylab("Visitors (per 1000 people)") +
  ggtitle("B") + # label the panel
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=10, family = "sans", face = "bold"),
        axis.title.x = element_text(size=10, family = "sans", face = "bold"),
        axis.text.y = element_text(size=7, family = "sans"),
        axis.text.x  = element_text(size=7, family = "sans"),
        axis.ticks.x = element_blank(),
        plot.title = element_text(size = 25, family = "sans", face = "bold",
                                  vjust = -5.5, hjust = 0.02),
        legend.position = "none",
        legend.box.background = element_rect(color = "black"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0,0.2,0,0.2), "cm"))
seasonal
```

### Figure 1C/D: Attendance Trends by Weather

In Figure 1, the relationship between attendance and temperature/precipitation were visualized for August and December. However, identical figures were produced for all months of the year and are accessible at [our GitHub repository](https://github.com/QuantitativeEcologyLab/BCParks_Attendance/tree/main/Figures/historical%20trends).

```{r fig1CD}
# Plot for August
august <-
  ggplot() +
  geom_point(data = bcparks[which(bcparks$month == "8"),], # point for every august observation
             aes(y = avgprecip, x = avgtemp, col = region),
             size = sqrt(bcparks[which(bcparks$month == "8"),]$attendance), # make point size proportional to attendance rate
             alpha = 0.5,
             shape = 16) + # point for every observation
  xlab("Temperature (ºC)") +
  ylab("Precipitation (mm)") +
  ggtitle("C") + # label the panel
  labs(subtitle = "August Attendance") +
  scale_y_continuous(limits = c(0,7)) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +  
  labs(size='Visitors per \n1000 people') +
  scale_size_continuous(range  = c(0.1, 8), 
                        limits = c(0, 10), 
                        breaks = c(2, 4, 6, 8, 10)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=10, family = "sans", face = "bold"),
        axis.title.x = element_text(size=10, family = "sans", face = "bold"),
        axis.text.y = element_text(size=7, family = "sans"),
        axis.text.x  = element_text(size=7, family = "sans"),
        plot.subtitle = element_text(size = 10, family = "sans", face = "bold"),
        plot.title = element_text(size = 25, family = "sans", face = "bold",
                                  vjust = -8, hjust = 0.02),
        legend.position = "none",
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0,0.2,0,0.2), "cm"))
august

# Plot for December
december <-
  ggplot() +
  geom_point(data = bcparks[which(bcparks$month == "12"),], # point for every december observation
             aes(y = avgprecip, x = avgtemp, col = region),
             size = sqrt(bcparks[which(bcparks$month == "12"),]$attendance), # make point size proportional to attendance rate
             alpha = 0.5,
             shape = 16) +
  xlab("Temperature (ºC)") +
  ylab("Precipitation (mm)") +
  ggtitle("D") + # label the panel
  labs(subtitle = "December Attendance") +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +  
  labs(size='Visitors per \n1000 people') +
  scale_size_continuous(range  = c(0.1, 8), 
                        limits = c(0, 10), 
                        breaks = c(2, 4, 6, 8, 10)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=10, family = "sans", face = "bold"),
        axis.title.x = element_text(size=10, family = "sans", face = "bold"),
        axis.text.y = element_text(size=7, family = "sans"),
        axis.text.x  = element_text(size=7, family = "sans"),
        plot.subtitle = element_text(size = 10, family = "sans", face = "bold"),
        plot.title = element_text(size = 25, family = "sans", face = "bold",
                                  vjust = -8, hjust = 0.02),
        legend.position = "none",
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0,0.2,0,0.2), "cm"))
december
```

```{r 1CD legend}
# Plot both August and December observations to ensure legend has common point sizes
plot <-
  ggplot() +
  geom_point(data = bcparks[which(bcparks$month == "8"),], 
             aes(y = avgprecip, x = avgtemp, 
                 size = sqrt(bcparks[which(bcparks$month == "8"),]$attendance)),
             # alpha = 0.2,
             shape = 16) +
  geom_point(data = bcparks[which(bcparks$month == "12"),], 
             aes(y = avgprecip, x = avgtemp, 
                 size = sqrt(bcparks[which(bcparks$month == "12"),]$attendance)),
             # alpha = 0.2,
             shape = 16) +
  labs(size='Visitors per \n1000 people') +
  scale_size_continuous(range  = c(0.1, 8), 
                        limits = c(0, 10), 
                        breaks = c(2, 4, 6, 8, 10),
                        labels = c("2" = "4", 
                                   "4" = "16", 
                                   "6" = "36",
                                   "8" = "64",
                                   "10" = "100")) + # change labels, since point size is sqrt transformed
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=10, family = "sans", face = "bold"),
        axis.title.x = element_text(size=10, family = "sans", face = "bold"),
        axis.text.y = element_text(size=7, family = "sans"),
        axis.text.x  = element_text(size=7, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 25, family = "sans", face = "bold"),
        legend.title = element_text(size = 10, face = "bold"),
        legend.text = element_text(size = 10),
        legend.box.background = element_rect(color = "black"),
        legend.position = "right",
        legend.direction = "horizontal",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0,0.2,0,0.2), "cm"))
```

### Combine all panels

```{r fig1, fig.height = 8, fig.width = 10}
# Make a function to grab legend from figures
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
# Save region legend from map
regionlegend <- get_legend(map)
# Save size legend from attendance/weather plot
sizelegend <- get_legend(plot)
# Remove legends from all plot
map <- map + theme(legend.position="none")
august <- august + theme(legend.position="none")
december <- december + theme(legend.position="none")
seasonal <- seasonal + theme(legend.position="none")

# Combine all 4 panels plus two legends
fig <- arrangeGrob(arrangeGrob(map, seasonal,
                               ncol = 1,
                               heights = c(8,6)), # left half
                   arrangeGrob(august, december, sizelegend,
                               ncol = 1,
                               heights = c(8,8,3)), # right half
                   ncol=2, widths = c(8, 6))
FIG1 <- grid.arrange(regionlegend, fig,
                     nrow = 2, heights = c(1,20))
```


## Recreate Figure 2

### Figure 2A: Attendance and Temperature

```{r fig2, fig.height = 5, fig.width = 10}
# Plot attendance in response to temperature
temp <-
  ggplot(data = bcparks[which(bcparks$attendancetype == "dayuse"),], 
         aes(y = attendance, x = avgtemp)) +
  geom_point(aes(col = region),
             alpha = 0.01) + # points for each observation
  geom_line(stat="smooth",method = "gam",
            aes(col = region),
            alpha = 0.9, size = 0.7, se = F) + # line for each region
  geom_line(stat="smooth",method = "gam",
            alpha = 1, col = "black", size = 1.2, se = T) + # line for overall trend
  scale_y_continuous(limits = c(0,10)) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +  
  xlab("Average Monthly Temperature (ºC)") +
  ylab("Monthly Park Visitors (per 1000 people)") +
  ggtitle("A") +
  # labs(subtitle = "Attendance and Temperature") +
  guides(col = guide_legend(override.aes = list(alpha=1))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=10, family = "sans"),
        axis.text.x  = element_text(size=8, family = "sans"),
        plot.title = element_text(hjust = 0.04, vjust = -5, size = 35, 
                                  family = "sans", face = "bold"),
        # plot.subtitle = element_text(size = 14, family = "sans", face = "bold"),
        legend.position=c(0.21,0.66),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(face = "bold"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))
temp
```

### Figure 2B: Attendance and Precipitation

```{r}
# Plot attendance in response to precipitation
precip <-
  ggplot(data = bcparks[which(bcparks$attendancetype == "dayuse"),], 
         aes(y = attendance, x = avgprecip)) +
  geom_point(aes(col = region),
             alpha = 0.01) + # points for each observation
  geom_line(stat="smooth",method = "gam",
            aes(col = region),
            alpha = 0.9, size = 0.7, se = F) + # line for each region
  geom_line(stat="smooth",method = "gam",
            alpha = 1, col = "black", size = 1.2, se = T) + # line for overall trend
  scale_y_continuous(limits = c(0,8)) +
  scale_x_continuous(limits = c(0,30)) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +  
  xlab("Average Monthly Precipitation (mm)") +
  ylab("Monthly Park Visitors (per 1000 people)") +
  ggtitle("B") +
  # labs(subtitle = "Attendance and Precipitation") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=10, family = "sans"),
        axis.text.x  = element_text(size=8, family = "sans"),
        plot.title = element_text(hjust = 0.04, vjust = -5, size = 35, 
                                  family = "sans", face = "bold"),
        # plot.subtitle = element_text(size = 10, family = "sans", face = "bold"),
        legend.position = "none",
        legend.box.background = element_rect(color = "black"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))
precip
```

### Combine the panels

```{r}
# Combine the 2 panels
FIG2 <- grid.arrange(temp, precip, 
                     ncol=2, widths = c(6,5.2))
```


# Modelling

## Build the HGAM

Since attendance was clearly influenced on weather and season, we included smooth terms of temperature, precipitation, and month in our model. In addition, we included interaction terms between month and temperature, month and precipitation, and temperature and precipitation. These interactions were meant to capture the correlation between month and weather, as well as the correlation between temperature and precipitation themselves.

We supplied our model with additional data from 2019 to 2022 for 3 Okanagan parks; Fintry Park, Skaha Bluffs Park, and E.C. Manning.

```{r model, eval = FALSE}
# Import additional data
ok2019_2022 <- read.csv("~/Desktop/bio/440/Okanagan/Data/ok2019-2022.csv")
# Add columns necessary to bind to `bcparks` dataset
ok2019_2022$region <- "ok"
ok2019_2022$attendancetype <- "dayuse"
ok2019_2022 <- ok2019_2022 %>%
  mutate(latitude = case_when(
    park == "Fintry Park" ~ 50.138676,
    park == "Skaha Bluffs Park" ~ 49.433939,
    park == "E.C. Manning Park" ~ 49.119421
  ))
ok2019_2022 <- ok2019_2022 %>%
  mutate(longitude = case_when(
    park == "Fintry Park" ~ -119.501595,
    park == "Skaha Bluffs Park" ~ -119.546110,
    park == "E.C. Manning Park" ~ -120.856781
  ))
# Add new rows to `bcparks` dataset: 2019-2022 data for 3 Okanagan Parks
bcparks <- rbind(bcparks,ok2019_2022) 

# Model with all day-use data (recent Okanagan data included)
M <- gam(attendance ~
           s(month, park, bs = 'fs', 
             xt = list(bs = 'cc'), # account for cyclic (annual) nature of attendance
             k = 5) +
           s(avgtemp) + 
           s(log(avgprecip + 1e-10)) + # add a tiny number to avgprecip so we don't take log of 0
           avgtemp:log(avgprecip + 1e-10) +
           month:avgtemp + 
           month:log(avgprecip + 1e-10), 
         family = Gamma(link = 'log'), # attendance can never be negative
         data = bcparks,
         method = 'REML',
         knots = list(month = c(0.5, 12.5))) # state that month 1 (Jan) follows month 12 (Dec)
```

```{r model summary}
summary(M)
```

## Check the Model

To ensure the behaviour of our model was appropriate based on observed data, we predicted monthly attendance rates for historical years (2010 through 2019). Then, we visually compared the predicted values to the historical counts for each month. We performed these comparisons for 2 random parks in each region.

```{r set seed, include = FALSE}
set.seed(42)
```


```{r model check}
#Build datasets of a couple random parks for each region
group <- bcparks %>% 
  group_by(park) %>%
  groups %>%
  unlist %>% 
  as.character
#NORTHERN REGION
north_parks <- bcparks[which(bcparks$region == "north"),] %>% 
  group_by(park) %>% 
  summarise() %>% 
  sample_n(2) %>% 
  mutate(unique_id=1:NROW(.))
northpredict <- bcparks[which(bcparks$region == "north"),] %>% 
  group_by(park)  %>% 
  right_join(north_parks, by=group) %>%
  group_by_(group)
#SOUTH COAST
south_parks <- bcparks[which(bcparks$region == "south"),] %>% 
  group_by(park) %>% 
  summarise() %>% 
  sample_n(2) %>% 
  mutate(unique_id=1:NROW(.))
southpredict <- bcparks[which(bcparks$region == "south"),] %>% 
  group_by(park)  %>% 
  right_join(south_parks, by=group) %>%
  group_by_(group)
#WEST COAST
west_parks <- bcparks[which(bcparks$region == "west"),] %>% 
  group_by(park) %>% 
  summarise() %>% 
  sample_n(2) %>% 
  mutate(unique_id=1:NROW(.))
westpredict <- bcparks[which(bcparks$region == "west"),] %>% 
  group_by(park)  %>% 
  right_join(west_parks, by=group) %>%
  group_by_(group)
#THOMPSON-CARIBOO
tc_parks <- bcparks[which(bcparks$region == "tc"),] %>% 
  group_by(park) %>% 
  summarise() %>% 
  sample_n(2) %>% 
  mutate(unique_id=1:NROW(.))
tcpredict <- bcparks[which(bcparks$region == "tc"),] %>% 
  group_by(park)  %>% 
  right_join(tc_parks, by=group) %>%
  group_by_(group)
#KOOTENAY-OKANAGAN
ok_parks <- bcparks[which(bcparks$region == "ok"),] %>% 
  group_by(park) %>% 
  summarise() %>% 
  sample_n(2) %>% 
  mutate(unique_id=1:NROW(.))
okpredict <- bcparks[which(bcparks$region == "ok"),] %>% 
  group_by(park)  %>% 
  right_join(ok_parks, by=group) %>%
  group_by_(group)

#"Predict" 2010-2019 attendance for these few parks
#NORTHERN REGION
northpredict$predict <- predict(M, newdata = northpredict, type = "response")
#SOUTH COAST
southpredict$predict <- predict(M, newdata = southpredict, type = "response")
#WEST COAST
westpredict$predict <- predict(M, newdata = westpredict, type = "response")
#THOMPSON-CARIBOO
tcpredict$predict <- predict(M, newdata = tcpredict, type = "response")
#KOOTENAY-OKANAGAN
okpredict$predict <- predict(M, newdata = okpredict, type = "response")

#Visually compare historical vs. model-predicted attendance to make sure models are behaving 
#NORTHERN REGION
ggplot(data = northpredict) +
  geom_boxplot(aes(x = month, y = attendance, 
                   group = cut_width(month, 1))) + # box plots for historical data
  geom_point(aes(x = month, y = predict), 
             size = 2, alpha = 0.5, col = "#CC6677") + # point for every year's prediction
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  xlab("Month") +
  ylab("Park visitors (per 1000 people)") +
  ggtitle("Northern Parks: Historical vs. Projected attendance") +
  facet_wrap(~ unique_id, labeller = as_labeller(c(
    `1` = "Random park 1",
    `2` = "Random park 2"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=9, family = "sans"),
        axis.text.x  = element_text(size=9, angle = 45, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = "right",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

#SOUTH COAST REGION
ggplot(data = southpredict) +
  geom_boxplot(aes(x = month, y = attendance, 
                   group = cut_width(month, 1))) + # box plots for historical data
  geom_point(aes(x = month, y = predict), 
             size = 2, alpha = 0.5, col = "#DDCC77") + # point for every year's prediction
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  xlab("Month") +
  ylab("Park visitors (per 1000 people)") +
  ggtitle("South Coast Parks: Historical vs. Projected attendance") +
  facet_wrap(~ unique_id, labeller = as_labeller(c(
    `1` = "Random park 1",
    `2` = "Random park 2"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=9, family = "sans"),
        axis.text.x  = element_text(size=9, angle = 45, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = "right",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

#WEST COAST REGION
ggplot(data = westpredict) +
  geom_boxplot(aes(x = month, y = attendance, 
                   group = cut_width(month, 1))) + # box plots for historical data
  geom_point(aes(x = month, y = predict), 
             size = 2, alpha = 0.5, col = "#88CCEE") + # point for every year's prediction
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  xlab("Month") +
  ylab("Park visitors (per 1000 people)") +
  ggtitle("West Coast Parks: Historical vs. Projected attendance") +
  facet_wrap(~ unique_id, labeller = as_labeller(c(
    `1` = "Random park 1",
    `2` = "Random park 2"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=9, family = "sans"),
        axis.text.x  = element_text(size=9, angle = 45, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = "right",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

#THOMPSON-CARIBOO REGION
ggplot(data = tcpredict) +
  geom_boxplot(aes(x = month, y = attendance, 
                   group = cut_width(month, 1))) + # box plots for historical data
  geom_point(aes(x = month, y = predict), 
             size = 2, alpha = 0.5, col = "#117733") + # point for every year's prediction
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  xlab("Month") +
  ylab("Park visitors (per 1000 people)") +
  ggtitle("Thompson-Cariboo Parks: Historical vs. Projected attendance") +
  facet_wrap(~ unique_id, labeller = as_labeller(c(
    `1` = "Random park 1",
    `2` = "Random park 2"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=9, family = "sans"),
        axis.text.x  = element_text(size=9, angle = 45, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = "right",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

#KOOTENAY-OKANAGAN REGION
ggplot(data = okpredict) +
  geom_boxplot(aes(x = month, y = attendance, 
                   group = cut_width(month, 1))) + # box plots for historical data
  geom_point(aes(x = month, y = predict), 
             size = 2, alpha = 0.5, col = "#332288") + # point for every year's prediction
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  xlab("Month") +
  ylab("Park visitors (per 1000 people)") +
  ggtitle("Kootenay-Okanagan Parks: Historical vs. Projected attendance") +
  facet_wrap(~ unique_id, labeller = as_labeller(c(
    `1` = "Random park 1",
    `2` = "Random park 2"))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=12, family = "sans", face = "bold"),
        axis.title.x = element_text(size=12, family = "sans", face = "bold"),
        axis.text.y = element_text(size=9, family = "sans"),
        axis.text.x  = element_text(size=9, angle = 45, family = "sans"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = "right",
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

```

# Projecting

To project BC Parks attendance out to 2100, we required information on the future of BC's climate and population.

## Climate Change Projections

Like the historical climate data, the climate change projections were downloaded to attain average monthly temperature (ºC) and precipitation (mm) at each park (Wang et al., 2016). We cleaned the downloaded data so as to remove irrelevant variables and convert monthly total precipitation to average monthly precipitation. Again, average monthly temperatures and precipitation were represented by the columns `avgtemp` and `avgprecip`, respectively.

```{r cc projs, eval = FALSE}
# Download Climate Projections
for(y in 2021:2100) {
  cat('Downloading ', y, '...\n', sep = '') # to track progress
  projClimateNA(
    file = 'parks-dem.csv',
    tFrame = 'M', # monthly averages
    exe = 'ClimateNA_v7.31.exe', # exe location (in working directory)
    scen = '8GCM', # 8GCMs_ensemble General Circulation Model
    ssp = c('S1', 'S2', 'S3', 'S5'), # Shared Socioeconomic Pathway scenarios
    years = as.character(y))
}

# Clean Climate Projections
climate <-
  # list all files, and import each of the CSVs
  map_dfr(
    list.files(
      'Data/climate-projections/parks-dem',
      full.names = TRUE,
      pattern = '@'),
    \(.fname) {
      readr::read_csv(.fname, col_types = '?') %>%
        # add a column of the file name
        mutate(file = .fname)
    }) %>%
  # extract scenario and year from file name column
  mutate(scenario =
           substr(file,
                  start = stri_locate_last(file, regex = '/')[1] + 1,
                  stop = stri_locate_first(file, regex = '@')[1] - 1),
         year = substr(file,
                       start = stri_locate_first(file, regex = '@')[1] + 1,
                       stop = nchar(file) - nchar('.csv'))) %>%
  # only keep relevant columns
  select(scenario, year, Latitude, Longitude, Elevation, Tave01, Tave02,
         Tave03, Tave04, Tave05, Tave06, Tave07, Tave08, Tave09, Tave10,
         Tave11, Tave12, PPT01, PPT02, PPT03, PPT04, PPT05, PPT06, PPT07,
         PPT08, PPT09, PPT10, PPT11, PPT12) %>%
  # pivot from wide to long format (only one column of precip and temp)
  pivot_longer(-c(scenario, year, Latitude, Longitude, Elevation),
               names_to = 'parameter', values_to = 'value') %>%
  # extract month and year out of parameters
  mutate(month = map_chr(parameter,
                         \(.chr) substr(.chr, nchar(.chr) - 1, nchar(.chr))),
         month = as.numeric(month),
         year = as.numeric(year),
         parameter = map_chr(parameter,
                             \(.chr) substr(.chr, 1, nchar(.chr) - 2))) %>%
  # pivot wider to make separate columns of temperature and precipitation
  pivot_wider(names_from = parameter, values_from = value) %>%
  # convert monthly total precipitation to average monthly precipitation
  mutate(first_day = as.Date(paste(year, month, '01', sep = '-')),
         next_month = if_else(month != '12', as.numeric(month + 1), 1),
         next_year = if_else(month != '12', year, year + 1),
         last_day = as.Date(paste(next_year, next_month, '01', sep = '-')),
         samples = as.numeric((last_day - first_day)),
         avgprecip = PPT / samples) %>% # convert to millimeters per day
  # change to names used in the models
  rename(avgtemp = Tave,
         latitude = Latitude,
         longitude = Longitude,
         elevation = Elevation,
         ssp = scenario) %>%
  # remove unnecessary columns from dataset
  select(-c(elevation, first_day, last_day, next_month, next_year, samples, tot_precip)) %>%
  # move month column to after the year one
  relocate(month, .after = year)

# Rename SSP levels for climate projections
climate$scenario <- recode_factor(climate$scenario, 
                                  "8GCMs_ensemble_ssp126" = "1-2.6", 
                                  "8GCMs_ensemble_ssp245" = "2-4.5",
                                  "8GCMs_ensemble_ssp370" = "3-7.0",
                                  "8GCMs_ensemble_ssp585" = "5-8.5")

```

## Population Projections

To predict numbers of visitors (rather than attendance as a proportion of the population), we considered the future of BC's population. Statistics Canada (2019) offered 9 different population growth scenarios; fast-aging, slow-aging, high growth, medium growth 1, medium growth 2, medium growth 3, medium growth 4, medium growth 5, and low growth. Under each of these scenarios, Statistics Canada (2019) predicted BC population growth rates for 2022/2023, 2032/2033, and 2042/2043. We projected growth rates between these decades and beyond to 2100 and used them to calculate the number of BC residents. This study focused on attendance under the high and low growth scenarios (HG and LG, respectively), but datasets with projected population under all scenarios are available on [our GitHub repository](https://github.com/QuantitativeEcologyLab/BCParks_Attendance/tree/main/Data/projections), as well as [a figure visualizing all growth rate projections](https://github.com/QuantitativeEcologyLab/BCParks_Attendance/blob/main/Figures/population%20growth/population-growth-rates.png).

```{r pop growth data, eval = FALSE}
# Download population growth rate file
popgrowth <- read.csv("~/Desktop/bio/440/BCParks_Attendance/Data/population/popgrowth.csv")

# Ensure correct formatting  
popgrowth$scenario <- as.factor(popgrowth$scenario)
```

### Project Growth Rates

```{r pop growth projs, eval = FALSE}
#Fit the model
popmodel <- glmer(growthrate ~ year  + (year|scenario),
                  family = Gamma(link="log"),
                  data = popgrowth)

summary(popmodel)

# Make dataset projecting LG growth rates using model
LGrates <- data.frame(year = seq(2022,2100, 0.1),
                      scenario = "LG")
LGrates$growthrate <- predict(popmodel, newdata = LGrates, type = "response")
LGrates$population <- 0 # make population column
LGrates[1,4] = LGrates[1,4] + 5319324 # set BC's current population in cell [1,4]
for(i in 2:nrow(LGrates)){
  LGrates[i,4] <- (LGrates[i,3]/1000*LGrates[i-1,4])+LGrates[i-1,4]
} # calculate population based on growth rates and BC's current population


# Make dataset projecting HG growth rates using model
HGrates <- data.frame(year = seq(2020,2100, 1),
                      scenario = "HG")
HGrates$growthrate <- predict(popmodel, newdata = HGrates, type = "response")
HGrates$population <- 0 # make population column
HGrates[1,4] = HGrates[1,4] + 5319324 # set BC's current population in cell [1,4]
for(i in 2:nrow(HGrates)){
  HGrates[i,4] <- (HGrates[i,3]/1000*HGrates[i-1,4])+HGrates[i-1,4]
} # calculate population based on growth rates and BC's current population

```

## Project Attendance Under Four SSPs and Two Population Growth Scenarios

We predicted attendance rates (visitors per 1,000 BC residents) under four climate change scenarios (SSP 1-2.6, SSP 2-4.5, SSP 3-7.0, and SSP 5-8.5). To calculate numbers of visitors from the proportions, we created separate datasets for the two population growth scenarios and performed identical calculations.

```{r projs, eval = FALSE}
# Model requires park column, so annotate coordinates with park names
climate <- merge(x=climate,y=coords, 
                 by=c("latitude","longitude"), all.x=TRUE)

# Predict attendance per 1,000 based on climate projections
climate$predicted_attendance <- predict(M, newdata = climate, type = "response")
# Ensure there are no NA values
climate <- na.omit(climate)

# Make datasets for different population growth scenarios
LGattendance <- merge(climate,LGrates,by=c("year"))
HGattendance <- merge(climate,HGrates,by=c("year"))

# Reorganize columns
LGattendance <- LGattendance %>% relocate(c(park, year, month, region, avgtemp, avgprecip, 
                                            predicted_attendance, latitude, longitude, elevation), .after = ssp)
HGattendance <- HGattendance %>% relocate(c(park, year, month, region, avgtemp, avgprecip, 
                                            predicted_attendance, latitude, longitude, elevation), .after = ssp)
# Make the SSP/year/month columns ascending & alphabetize the park column
LGattendance <- LGattendance[order(LGattendance$ssp, LGattendance$park, LGattendance$year, LGattendance$month),]
HGattendance <- HGattendance[order(HGattendance$ssp, HGattendance$park, HGattendance$year, HGattendance$month),]

# Calculate projected visitor COUNTS for each population growth scenario
LGattendance$predicted_visitors <- (LGattendance$predicted_attendance/1000)*LGattendance$population
HGattendance$predicted_visitors <- (HGattendance$predicted_attendance/1000)*HGattendance$population

# Ensure correct formatting of predictions
LGattendance$predicted_visitors <- as.numeric(LGattendance$predicted_visitors)
HGattendance$predicted_visitors <- as.numeric(HGattendance$predicted_visitors)
```

# Visualizing Results

## Recreate Figure 3

### Change in Annual Visitors Relative to 2020

As these parks range vastly in attendance rates, we calculated a change in annual visitors relative to the first year of predictions (2020) for easy visualization.

```{r relative visitors}
# Calculate change in attendance relative to first prediction
LGattendance <- LGattendance %>%
  group_by(ssp, year, park) %>% # don't compare different SSPs, years, or parks to each other
  mutate(annual_visitors = sum(predicted_visitors)) # get annual totals
LGattendance <- LGattendance %>%
  group_by(park, ssp) %>% # don't compare different SSPs or parks to each other
  mutate(relative_visitors = annual_visitors / first(annual_visitors)) 

HGattendance <- HGattendance %>%
  group_by(ssp, year, park) %>% # don't compare different SSPs, years, or parks to each other
  mutate(annual_visitors = sum(predicted_visitors)) # get annual totals
HGattendance <- HGattendance %>%
  group_by(park, ssp) %>% # don't compare different SSPs or parks to each other
  mutate(relative_visitors = annual_visitors / first(annual_visitors)) 

```

### Plot

```{r fig3, fig.height = 5.53, fig.width = 8.21}
# Make colour strips in x-direction for panel title boxes (they will correspond to SSPs)
strip <- strip_themed(background_x = 
                        elem_list_rect(fill = c("#4EBAF9", "#C0DEED", "#FFC1B5", "#FF7B7B")))

# Assign a letter for each SSP so we can label each panel
data_text <- data.frame(label = c("A", "B", "C", "D"),
                        ssp = names(table(LGattendance$ssp)),
                        x = c(2023, 2023, 2023, 2023), 
                        y = c(1.9, 1.9, 1.9, 1.9))

# Plot relative growth in visitors (same values for both LG and HG, so pick either dataset)
projections <-
  ggplot(LGattendance, aes(x = year, y = relative_visitors)) + 
  geom_hline(yintercept = 1, size = 0.5, color = "grey70") + # line at 1 for reference
  geom_line(aes(group = park, col = region), 
            size=0.5, alpha = 0.1) + # line for each park
  facet_wrap2(~ ssp, strip = strip, labeller = as_labeller(c(
    `1-2.6` = "SSP 1-2.6",
    `2-4.5` = "SSP 2-4.5",
    `3-7.0` = "SSP 3-7.0",
    `5-8.5` = "SSP 5-8.5"))) + # create a panel for each climate change scenario
  geom_text(data = data_text,
            mapping = aes(x = x,
                          y = y,
                          label = label),
            check_overlap = TRUE,
            size = 8, fontface = "bold") + # label each panel
  xlab("Year") +
  ylab("Relative Change in Annual Visitors") +
  scale_y_continuous(limits = c(0.75,2)) +
  guides(col = guide_legend(override.aes = list(alpha=1))) +
  scale_colour_muted(name="Region",
                     labels=c('Northern', 
                              'Kootenay-Okanagan', 
                              'South Coast', 
                              'Thompson-Cariboo', 
                              'West Coast')) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.y = element_text(size=14, family = "sans", face = "bold"),
        axis.title.x = element_text(size=14, family = "sans", face = "bold"),
        axis.text.y = element_text(size=10, family = "sans"),
        axis.text.x  = element_text(size=10, family = "sans"),
        plot.title = element_text(size = 14, family = "sans", face = "bold"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.position=c(0.4,0.86),
        legend.key.size = unit(0.4, 'cm'),
        legend.title = element_text(size = 8, face = "bold"),
        legend.text=element_text(size=7),
        legend.box.background = element_rect(color = "black"),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))
projections
```

## Recreate Figure 4

### Case Study: Golden Ears Park

To closely examine the model's predictions for seasonal trends under different population growth scenarios, we visualized the results of a single park; Golden Ears Park. Identical figures for all parks are available on [our GitHub repository](https://github.com/QuantitativeEcologyLab/BCParks_Attendance/tree/main/Figures/seasonal%20projections).

```{r fig4, fig.height = 6, fig.width = 10}
# Disable scientific notation (for y-axis)
options(scipen = 999)

# Plot attendance under low population growth
LG <-
  ggplot() +
  facet_wrap2(~ ssp, strip = strip, labeller = as_labeller(c(
    `1-2.6` = "SSP 1-2.6",
    `2-4.5` = "SSP 2-4.5",
    `3-7.0` = "SSP 3-7.0",
    `5-8.5` = "SSP 5-8.5"))) +
  geom_line(data = LGattendance[which(LGattendance$park == "Golden Ears Park"),], 
            aes(month, predicted_visitors, group = year, col = year), 
            size=0.1) + # lines for each year
  geom_point(data = bcparks[which(bcparks$park == "Golden Ears Park"),] %>%
               group_by(month) %>%
               summarise(visitortotal = mean(visitortotal)), 
             aes(month, visitortotal), 
             col = "black", size = 1, na.rm = TRUE) + # plot historical monthly averages as points
  xlab("Month") +
  ylab("Monthly Visitors") +
  ggtitle("A") +
  labs(subtitle = "Golden Ears Park Attendance under Low Population Growth") +
  scale_y_continuous(labels = scales::comma, limits = c(0,500000)) +
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  scale_colour_viridis_c(name = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.y = element_text(size=14, family = "sans", face = "bold"),
        axis.title.x = element_text(size=14, family = "sans", face = "bold"),
        axis.text.y = element_text(size=10, family = "sans"),
        axis.text.x  = element_text(size=8, family = "sans", angle = 45, hjust = 1),
        plot.subtitle = element_text(size = 10, family = "sans", face = "bold"),
        plot.title = element_text(size = 25, family = "sans", face = "bold"),
        legend.position=c(0.11,0.84),
        legend.key.height = unit(0.4, 'cm'),
        legend.key.width = unit(0.6, 'cm'),
        legend.key.size = unit(0.25, 'cm'),
        legend.title = element_text(size = 8, family = "sans", face = "bold"),
        legend.text=element_text(size=7),
        legend.box.background = element_rect(color = "black"),
        legend.margin=margin(c(8,8,8,8)),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))
# Plot attendance under high population growth
HG <-
  ggplot() +
  facet_wrap2(~ ssp, strip = strip, labeller = as_labeller(c(
    `1-2.6` = "SSP 1-2.6",
    `2-4.5` = "SSP 2-4.5",
    `3-7.0` = "SSP 3-7.0",
    `5-8.5` = "SSP 5-8.5"))) +
  geom_line(data = HGattendance[which(HGattendance$park == "Golden Ears Park"),], 
            aes(month, predicted_visitors, group = year, col = year), 
            size=0.1) + # lines for each year
  geom_point(data = bcparks[which(bcparks$park == "Golden Ears Park"),] %>%
               group_by(month) %>%
               summarise(visitortotal = mean(visitortotal)), 
             aes(month, visitortotal), 
             col = "black", size = 1, na.rm = TRUE) + # plot historical monthly averages as points
  xlab("Month") +
  ylab("Monthly Visitors") +
  ggtitle("B") +
  labs(subtitle = "Golden Ears Park Attendance under High Population Growth") +
  scale_y_continuous(labels = scales::comma, limits = c(0,500000)) +
  scale_x_continuous(breaks = seq_along(month.abb), labels = month.abb) +
  scale_colour_viridis_c(name = "Year") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size=14, family = "sans", face = "bold"),
        axis.text.y = element_blank(),
        axis.text.x  = element_text(size=8, family = "sans", angle = 45, hjust = 1),
        plot.subtitle = element_text(size = 10, family = "sans", face = "bold"),
        plot.title = element_text(size = 25, family = "sans", face = "bold"),
        legend.position = "none",
        legend.title = element_text(size = 12, family = "sans"),
        legend.box.background = element_rect(color = "black"),
        legend.margin=margin(c(8,8,8,8)),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))

# Combine the 2 panels
FIG4 <- grid.arrange(LG, HG, 
                     ncol=2, widths = c(6,5.2))

```

# Session Info

```{r sessionInfo}
sessionInfo()
```
